Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control: A Secondary Analysis of SPRINT

### 2022-03-18 doc/manuscript-rmd.Rmd
draft rdy for NMP review; time to submit?

__RESULTS__. Among `r n_total` randomized participants, the mean (standard deviation) age was `r inline_text(tbl_characteristics, variable = 'age_yrs', column = 'stat_0_1')` years and `r inline_text(tbl_characteristics, variable = 'sex', column = 'stat_0_1')`% were women. Over the median intervention phase of 3.3 years, intensive treatment was beneficial for both cardiovascular mortality (Hazard Ratio [HR] = `r with(ap_l_o_t, .("{cvd_est}, 95% confidence interval [CI] {cvd_lwr} to {cvd_upr}"))`) and all-cause mortality (HR = `r with(ap_l_o_t, .("{mort_est}, 95% CI {mort_lwr} to {mort_upr}"))`). However, at the median total follow-up of `r ap_b_o$fup_est` years, there was no longer evidence of benefit for cardiovascular mortality (HR = `r with(ap_l_o_c, .("{cvd_est}, 95% CI {cvd_lwr} to {cvd_upr}"))`) or all-cause mortality (HR = `r with(ap_l_o_c, .("{mort_est}, 95% CI {mort_lwr} to {mort_upr}"))`). In a subgroup of `r n_ehr` participants, the estimated mean (95% CI) outpatient SBP among participants randomized to intensive treatment increased from `r f(bp_viz_5yr$Intensive)` at five years to `r f(bp_viz_10yr$Intensive)` ten years post-randomization.

@@

__Trial Design__: The design and methods of SPRINT have been published previously [@ambrosius_design_2014; @sprintFinal]. Briefly, it was a multicenter randomized clinical trial that compared two strategies for managing SBP in older adults with hypertension who were at increased risk of CVD. Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (CKD), or a Framingham Risk Score of 15% or greater or if they were aged 75 years or older. Individuals residing in a nursing home, persons with a diagnosis of dementia (based on medical record review), and those treated with medications primarily used for dementia therapy were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. Randomization began on November 8, 2010 and ended in March 2013. In addition to the trial’s intervention phase, which spanned November 8, 2010 through August 20, 2015, when the administrative decision to stop the trial was made, the current study includes as part of the “trial phase” the additional period of study-provided antihypertensive medications prior to the final closeout visit on July 1, 2016 (eFigure 1). Observational follow-up continued through December 2020

As trial follow-up visits and provision of antihypertensive medication continued through July 2016 [@sprintFinal], the current study considers a trial phase spanning from November 8, 2010 through July 2016 and an observational phase spanning from August 2017 through December, 2020 (__eFigure 1__). The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).

__Baseline Study Measurements__: Sociodemographic data were collected at baseline, with race or ethnicity information collected via self-report. The estimated glomerular filtration rate (eGFR) was calculated by the race-free 2021 CKD-EPI creatinine equation [@inker_new_2021]. Cognitive function was assessed using the Montreal Cognitive Assessment (MoCA) [@nasreddine_montreal_2005]. Lower cognitive function was defined as scoring at or below the estimated age and education-specific normative 10th percentile from the Irish Longitudinal Study of Aging [@kenny_normative_2013], after adding three points to the scores of non-White participants [@sachsMoCA2021]. We defined frailty status at baseline using a 36-item Frailty Index (FI) based upon the model of deficit accumulation [@pajewski_characterizing_2016]. The FI is calculated as the sum of the score for each deficit divided by the total number of nonmissing items. We categorized frailty status as fit (FI ≤ 0.10), less fit (0.10 < FI ≤ 0.21), or frail (FI > 0.21).

__National Death Index Linkage__: Outcomes of interest included all-cause and cardiovascular mortality. Methods of ascertainment and adjudication through the course of trial follow-up have been previously described [@sprintFinal]. In the final report of the SPRINT trial, mortality was ascertained through a US National Death Index (NDI) search completed in December 2016 [@sprintFinal]. For the current analysis, we completed an NDI search including deaths through 2020. As the NDI is continuously updated, the current analysis includes some deaths in 2016 that were not identified at the time of the final SPRINT report. Possible matches were identified according to NDI guidelines [@ndi_guide]. Deaths were treated as confirmed if they were a Class 1 match, or a Class 2, 3, or 4 match with a probabilistic score above cutoffs recommended by the NDI [@ndi_guide]. Deaths ascertained in 2020 were based on the NDI preliminary data release. CVD mortality for NDI-based follow-up used the NDI Plus System, which automatically identifies underlying causes of death from death certificates, including conversion to ICD-10 codes. We defined CVD mortality as any death containing the ICD-10 codes of I00 to I99 [@regards_ndi].

__EHR Ancillary Study__: Methods for the linkage of participants to their medical record number and the extraction of vital sign data have been previously described [@drawz_concordance_2020]. We identified 3074 participants with 3 or more outpatient and trial BP measurements. After excluding `r 3074 - n_ehr` participants without EHR data following July 2016 (i.e., conclusion of the trial phase), a total of `r n_ehr` patients were included for the ancillary BP analysis. Because encounter type information was inconsistently available (i.e. outpatient, inpatient, observation, etc.), we defined a BP measurement as outpatient if there was not a BP measurement on the preceding or following day, and if there were 2 or less BP measurements on a particular day. We averaged outpatient EHR BP readings when there were 2 on the same day.

@@

We examined the trajectory of SBP following the conclusion of the trial using outpatient SBPs extracted from the EHR. Mean between-group differences in outpatient SBP were estimated using linear mixed models. Models included random effects for participant and clinic site and an interaction between treatment group and time since randomization, which was modeled using B-splines. All analyses were performed using SAS version 9.4 (SAS Institute Inc, Cary, NC) and R version `r paste(version$major, version$minor, sep = '.')` (R Project for Statistical Computing [http://www.r-project.org]) with multiple auxiliary R packages.[@r_language; @table.glue; @tidyverse; @survival_package; @timereg; @targets] All hypothesis tests were 2-sided, and P values less than 0.05 were considered statistically significant. No adjustments for multiple comparisons were made.

@@

___Study Participants___. A total of `r n_total` participants were randomized between November 2010 and March 2013. The mean (standard deviation [SD]) age was `r inline_text(tbl_characteristics, variable = 'age_yrs', column = 'stat_0_1')` years, with `r inline_text(tbl_characteristics, variable = 'age_cat', column = 'stat_0_1')`% of participants aged 75 years or older (__eTable 1__). Participants were `r inline_text(tbl_characteristics, variable = 'sex', column = 'stat_0_1')`% female and `r inline_text(tbl_characteristics, variable = 'race', column = 'stat_0_1')`% black. The mean (SD) SBP at baseline was `r inline_text(tbl_characteristics, variable = 'sbp', column = 'stat_0_1')` and `r inline_text(tbl_characteristics, variable = 'moca', column = 'stat_0_1')`% of participants MoCA scores were below the 10^th^ percentile. Compared to participants not included in the ancillary EHR study, participants included were more likely to be male, were older, had lower SBP and higher scores on the MoCA, and had a higher prevalence of CKD.

@@

___All-cause Mortality___. In both treatment groups, the total median follow-up time was `r ap_b_o$fup_est` years. A total of `r ap_b_i$n_acm` and `r ap_b_s$n_acm` deaths occurred among participants randomized to intensive and standard treatment, respectively (__eTable 2__). The hazard ratio (HR) for all-cause mortality comparing intensive to standard treatment was `r with(ap_l_o_t, .("{mort_est} (95% confidence interval [CI] {mort_lwr}, {mort_upr})"))` during the trial phase, and `r with(ap_l_o_c, .("{mort_est} (95% CI {mort_lwr}, {mort_upr})"))` during the observational phase. The continuous time-dependent effect of intensive versus standard treatment indicated a benefit for all-cause mortality from `r t_signif_acm` to `r t_atten_acm` years from randomization, and was attenuated throughout the remainder of the observational phase (__Figure 1__).  In subgroups based on age, sex, race, CKD, cognitive function, and frailty status, there was no evidence that intensive treatment during the trial phase produced benefit for all-cause mortality during the observational phase of follow-up (__Figure 2__).

@@

___Cardiovascular Mortality___. A total of `r ap_b_i$n_cvd` and `r ap_b_s$n_cvd` CVD deaths occurred among participants randomized to intensive and standard treatment, respectively (__eTable 3__). The HR for CVD mortality among participants randomized to intensive versus standard treatment was `r with(ap_l_o_t, .("{cvd_est} (95% CI {cvd_lwr}, {cvd_upr})"))` during the trial phase and `r with(ap_l_o_c, .("{cvd_est} (95% CI {cvd_lwr}, {cvd_upr})"))` during the observational phase. The time-dependent effect of intensive versus standard treatment indicated a benefit for CVD mortality from `r t_signif_cvd` to `r t_atten_cvd` years from randomization, and was attenuated throughout the remainder of the observational phase (__Figure 3__). Adults randomized to intensive treatment who were <75 years of age, men, non-black, without CKD, or with cognitive function >10^th^ percentile had lower CVD mortality risk during the trial phase compared to their counterparts randomized to standard treatment, but there was no evidence that intensive treatment during the trial phase produced benefit for CVD mortality during the observational phase (__Figure 4__).

___Blood Pressure___. Amongst a subset of `r n_ehr` trial participants, the median number of outpatient BP measurements extracted from the EHR during the observational phase of follow-up was 20 (IQR 10 to 34). The estimated mean (95% CI) SBP among participants randomized to intensive treatment was `r f(bp_viz_5yr$Intensive)` at 5 years and `r f(bp_viz_10yr$Intensive)` at 10 years post-randomization (__Figure 5__). For participants randomized to standard treatment, mean (95% CI) SBP was estimated to be `r f(bp_viz_5yr$Standard)` at 5 years and `r f(bp_viz_10yr$Standard)` post-randomization. The between-group difference in mean SBP levels (intensive minus standard)  was `r f(bp_diff_5yr)` mm Hg at 5 years post-randomization and was reduced to `r f(bp_diff_5yr)` at 10 years post-randomization (__eFigure 2__).

@@

The current study analyzed all-cause and CVD-mortality among SPRINT participants up to 10 years following randomization, finding that the benefits associated with intensive treatment quickly attenuated as SBP levels increased after the trial intervention was discontinued. Time-varying estimates of the benefit of intensive treatment for all-cause mortality were attenuated at `r t_atten_acm` years while the benefit for CVD-mortality was attenuated at `r t_atten_cvd` years post-randomization. Findings from our ancillary study of outpatient SBP measured in routine clinical practice indicated that the difference in SBP between treatment groups diminished over time, with no detectable difference in SBP approximately `r t_atten_sbp` years after randomization. These results in combination with the primary findings of SPRINT indicate that the beneficial effect of intensive treatment among adults with hypertension appears to diminish if BP control is not sustained.

The Strategy of Blood Pressure Intervention in the Elderly Hypertensive Patients (STEP) trial enrolled 8511 Chinese patients 60 to 80 years of age with hypertension and randomized patients to a SBP target of 110 to less than 130 mm Hg (intensive treatment) or a target of 130 to less than 150 mm Hg (standard treatment) [@zhang_trial_2021]. STEP found a HR of 0.72 (95% CI 0.39, 1.32) with intensive versus standard treatment for CVD mortality after a median follow-up of 3.34 years, but did not find evidence of a benefit for all-cause mortality. In the current study of SPRINT participants, the protective effect of intensive treatment for all-cause mortality was attenuated several years before the protective effect for CVD. These results in combination with findings from the STEP trial suggest weaker evidence for reduced all-cause versus CVD mortality risk with intensive BP control.

<!-- I commenting this sentence out because I think it may be too speculative for this Discussion. Nick, if you like the idea of including the commented sentence in the paper, can we set up a call with Paul Whelton to discuss? -->

<!-- Another contrasting result was that intensive treatment in SPRINT, but not STEP, increased the risk of acute kidney injury (AKI). While this may be attributable to the much lower prevalence of chronic kidney disease in STEP (2.3% vs 28.3%), the association of AKI with increased mortality risk may have been a contributing factor to the longer term effect of randomization to intensive treatment on all-cause mortality among SPRINT participants [@chawla2014aki; @rocco2018aki]. -->

<!-- Meta analyses of randomized trials have shown that pharmacological blood pressure (BP) lowering reduces the risk of major CVD events across the spectrum of initial BP [@rahimi_age_stratified_2021; @bundy_systolic_2017]. -->

Previous studies have found rising BP levels among US adults during the time period of the current study. General population studies of adults living in the US with hypertension found that the prevalence of uncontrolled BP (SBP ≥ 140 mm Hg or DBP ≥ 90 mm Hg) increased from 2013 to 2017 [@muntner2020htn]. In addition, an analysis of data from 464,585 adults enrolled in a Quest Diagnostics wellness program found SBP was 1 mm Hg and 3 mm Hg higher, depending on age group and sex, in April through December of 2020 versus their corresponding values throughout 2019 [@laffin_rise_2022]. The current study shows that even for adults who have maintained intense SBP control for 3 years, increasing SBP levels can quickly diminish the protective effect. Combined with previous findings on rising BP levels among US adults, data from the current study emphasize the need for implementation of population- and community-level strategies to improve BP control in the US.

@@

This study has several limitations. First, while we restricted analyses to high quality NDI matches, misclassification in linking participants to the NDI is possible. Second, while several studies have shown reasonable performance of using NDI diagnosis codes for defining CVD mortality [@regards_ndi], it is not as robust as the adjudication process used in the primary follow-up for the trial. Third, information about SBP control after the trial was limited to routine outpatient SBP values extracted from the EHR, which are known to poorly reflect the standardized BP measurement process used during the SPRINT trial [@drawz_concordance_2020]. While this prohibits definite conclusions about the absolute level of BP in both treatment groups, as well as pin-pointing when the between-group difference may have completely attenuated, the observation of steadily increasing SBP for participants in the intensive treatment and relatively stable SBP in the standard treatment group following the trial is likely still valid.

In conclusion, while intensive treatment produced beneficial effects on mortality during the trial, there was no evidence that this produced sustained benefits on cardiovascular and all-cause mortality subsequent to discontinuing the intervention protocol. Given steadily increasing mean SBP levels in participants randomized to intensive treatment after the trial, these results demonstrate that maintaining more intensive BP targets throughout adulthood will likely be essential for long-term CVD risk management.

>>>>

__RESULTS__. Among `r n_total` randomized participants, the mean (standard deviation) age was `r inline_text(tbl_characteristics, variable = 'age_yrs', column = 'stat_0_1')` years and `r inline_text(tbl_characteristics, variable = 'sex', column = 'stat_0_1')`% were women. Over the median intervention phase of 3.3 years, intensive treatment was beneficial for both cardiovascular mortality (Hazard Ratio [HR] = `r with(ap_l_o_t, .("{cvd_est}, 95% confidence interval [CI] {cvd_lwr} to {cvd_upr}"))`) and all-cause mortality (HR = `r with(ap_l_o_t, .("{mort_est}, 95% CI {mort_lwr} to {mort_upr}"))`). However, at the median total follow-up of `r ap_b_o$fup_est` years, there was no longer evidence of benefit for cardiovascular mortality (HR = `r with(ap_l_o_c, .("{cvd_est}, 95% CI {cvd_lwr} to {cvd_upr}"))`) or all-cause mortality (HR = `r with(ap_l_o_c, .("{mort_est}, 95% CI {mort_lwr} to {mort_upr}"))`). In a subgroup of `r n_ehr` participants, the estimated mean (95% CI) outpatient SBP among participants randomized to intensive treatment increased from `r f(bp_viz_5yr$Intensive)` at five years to `r f(bp_viz_10yr$Intensive)` ten years following randomization.

@@

__Trial Design__: The design and methods of SPRINT have been published previously [@ambrosius_design_2014; @sprintFinal]. Briefly, it was a multicenter randomized clinical trial that compared two strategies for managing SBP in older adults with hypertension who were at increased risk for CVD. Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the use and number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (CKD), a Framingham Risk Score of 15% or greater, or if they were aged 75 years or older. Individuals residing in a nursing home, with a diagnosis of dementia (based on medical record review), and those treated with medications prescribed for dementia were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. Randomization began on November 8, 2010 and ended in March 2013. On August 20, 2015, the Director of the National Heart, Lung, and Blood Institute accepted the data and safety monitoring board’s recommendation to inform the investigators and participants of the cardiovascular results, and decided to stop the trial, early, for benefit. In addition to the trial’s intervention phase, which spanned November 8, 2010 through August 20, 2015, when the administrative decision to stop the trial was made, the current study includes as part of the “trial phase” the additional period of study-provided antihypertensive medications prior to the final closeout visit on July 1, 2016 (eFigure 1). Observational follow-up continued through December 2020. The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).

__Baseline Study Measurements__: Sociodemographic data were collected at baseline, with race or ethnicity information collected via self-report. In the current study, the estimated glomerular filtration rate (eGFR) was calculated by the race-free 2021 CKD-EPI creatinine equation [@inker_new_2021]. Cognitive function was assessed using the Montreal Cognitive Assessment (MoCA) [@nasreddine_montreal_2005]. Lower cognitive function was defined as scoring at or below the estimated age and education-specific normative 10th percentile from the Irish Longitudinal Study of Aging [@kenny_normative_2013], after adding three points to the scores of non-White participants [@sachsMoCA2021]. We defined frailty status at baseline using a 36-item Frailty Index (FI) based upon the model of deficit accumulation [@pajewski_characterizing_2016]. The FI is calculated as the sum of the score for each deficit divided by the total number of nonmissing items. We categorized frailty status as fit (FI ≤ 0.10), less fit (0.10 < FI ≤ 0.21), or frail (FI > 0.21).

__National Death Index Linkage__: Outcomes of interest included all-cause and cardiovascular mortality. Methods of ascertainment and adjudication through the course of trial follow-up have been previously described [@sprintFinal]. In the final report of the SPRINT trial, mortality was ascertained through a US National Death Index (NDI) search completed in December 2016 [@sprintFinal]. For the current analysis, we completed an NDI search including deaths through December 2020. Possible matches were identified according to NDI guidelines [@ndi_guide]. Deaths were treated as confirmed if they were a Class 1 match, or a Class 2, 3, or 4 match with a probabilistic score above cutoffs recommended by the NDI [@ndi_guide]. Deaths ascertained in 2020 were based on the NDI preliminary data release. CVD mortality for NDI-based follow-up used the NDI Plus System, which automatically identifies underlying causes of death from death certificates, including conversion to ICD-10 codes. We defined CVD mortality as any death containing the ICD-10 codes of I00 to I99 [@regards_ndi].

__EHR Ancillary Study__: Methods for the linkage of participants to their medical record number and the extraction of vital sign data have been previously described [@drawz_concordance_2020]. We identified 3074 participants with 3 or more electronic health record reports of outpatient BP measurements during the trial. After excluding `r 3074 - n_ehr` participants without EHR data following July 2016 (i.e., conclusion of the trial phase), a total of `r n_ehr` patients were included for the ancillary BP analysis. Because encounter type information was inconsistently available (i.e. outpatient, inpatient, observation, etc.), we defined a BP measurement as outpatient if there was not a BP measurement on the preceding or following day, and if there were 2 or less BP measurements on a particular day. We averaged outpatient EHR BP readings when there were 2 on the same day.

@@

We examined the trajectory of SBP following the conclusion of the trial using outpatient SBPs extracted from the EHR. Mean between-group differences in outpatient SBP were estimated using linear mixed models. Models included random effects for participant and clinic site and an interaction between treatment group and time since randomization, which was modeled using B-splines. All analyses were performed using SAS version 9.4 (SAS Institute Inc, Cary, NC) and R version `r paste(version$major, version$minor, sep = '.')` (R Project for Statistical Computing [http://www.r-project.org]) with assistance from multiple R packages.[@r_language; @table.glue; @tidyverse; @survival_package; @timereg; @targets] All hypothesis tests were 2-sided, and P values less than 0.05 were considered statistically significant. No adjustments for multiple comparisons were made.

@@

___Study Participants___. A total of `r n_total` participants were randomized between November 2010 and March 2013. The mean (standard deviation [SD]) age was `r inline_text(tbl_characteristics, variable = 'age_yrs', column = 'stat_0_1')` years, with `r inline_text(tbl_characteristics, variable = 'age_cat', column = 'stat_0_1')`% of participants aged 75 years or older (__eTable 1__). Participants were `r inline_text(tbl_characteristics, variable = 'sex', column = 'stat_0_1')`% female and `r inline_text(tbl_characteristics, variable = 'race', column = 'stat_0_1')`% black. The mean (SD) SBP at baseline was `r inline_text(tbl_characteristics, variable = 'sbp', column = 'stat_0_1')` and `r inline_text(tbl_characteristics, variable = 'moca', column = 'stat_0_1')`% of participants MoCA scores were below the 10^th^ percentile. Compared to participants not included in the ancillary EHR study, participants included were more likely to be male, older, with lower SBP and higher scores on the MoCA, and a higher prevalence of CKD.

@@

___All-cause Mortality___. In both treatment groups, the total median follow-up time was `r ap_b_o$fup_est` years. A total of `r ap_b_i$n_acm` and `r ap_b_s$n_acm` deaths occurred among participants randomized to intensive and standard treatment, respectively (__eTable 2__). The hazard ratio (HR) for all-cause mortality comparing intensive to standard treatment was `r with(ap_l_o_t, .("{mort_est} (95% confidence interval [CI] {mort_lwr}, {mort_upr})"))` during the trial phase, and `r with(ap_l_o_c, .("{mort_est} (95% CI {mort_lwr}, {mort_upr})"))` during the observational phase. The continuous time-dependent effect of intensive versus standard treatment indicated a benefit for all-cause mortality from `r t_signif_acm` to `r t_atten_acm` years following randomization, and was attenuated throughout the remainder of the observational phase (__Figure 1__).  In subgroups based on age, sex, race, CKD, cognitive function, and frailty status, there was no evidence that intensive treatment during the trial phase produced benefit for all-cause mortality during the observational phase of follow-up (__Figure 2__).

@@

___Cardiovascular Mortality___. During the `r with(ap_b_o, table_glue("{fup_est}"))` years of follow-up, a total of `r ap_b_i$n_cvd` and `r ap_b_s$n_cvd` CVD deaths occurred among participants randomized to intensive and standard treatment, respectively (__eTable 3__). The HR for CVD mortality among participants randomized to intensive versus standard treatment was `r with(ap_l_o_t, .("{cvd_est} (95% CI {cvd_lwr}, {cvd_upr})"))` during the trial phase and `r with(ap_l_o_c, .("{cvd_est} (95% CI {cvd_lwr}, {cvd_upr})"))` during the observational phase. The time-dependent effect of intensive versus standard treatment indicated a benefit for CVD mortality from `r t_signif_cvd` to `r t_atten_cvd` years from randomization, and was attenuated throughout the remainder of the observational phase (__Figure 3__). Adults randomized to intensive treatment who were <75 years of age, men, non-black, without CKD, or with cognitive function >10^th^ percentile had a lower CVD mortality risk during the trial phase compared to their counterparts randomized to standard treatment, but there was no evidence that intensive treatment during the trial phase produced benefit for CVD mortality during the observational phase (__Figure 4__).

___Blood Pressure___. Amongst a subset of `r n_ehr` trial participants, the median number of outpatient BP measurements extracted from the EHR during the observational phase of follow-up was 20 (interquartile range: 10 to 34). The estimated mean (95% CI) SBP among participants randomized to intensive treatment was `r f(bp_viz_5yr$Intensive)` at 5 years following randomization and `r f(bp_viz_10yr$Intensive)` at 10 years following randomization (__Figure 5__). For participants randomized to standard treatment, mean (95% CI) SBP was estimated to be `r f(bp_viz_5yr$Standard)` at 5 years following randomization and `r f(bp_viz_10yr$Standard)` at 10 years following randomization. The between-group difference in mean SBP levels (intensive minus standard)  was `r f(bp_diff_5yr)` mm Hg at 5 years following randomization and was reduced to `r f(bp_diff_10yr)` at 10 years following randomization (__eFigure 2__).

@@

The current study analyzed all-cause and CVD-mortality among trial participants up to 10 years following randomization, finding that the benefits associated with intensive treatment quickly attenuated after the trial intervention was discontinued. Time-varying estimates of the benefit of intensive treatment for all-cause mortality were attenuated at `r t_atten_acm` years while the benefit for CVD-mortality was attenuated at `r t_atten_cvd` years following randomization. Findings from our ancillary study of outpatient SBP measured in routine clinical practice indicated that the difference in SBP between treatment groups diminished steadily over time, with no detectable difference in SBP approximately 9 years after randomization. These results in combination with the primary findings of the trial indicate that the beneficial effect of intensive treatment among adults with hypertension appears to diminish if BP control is not sustained.

The current study's results in the "trial phase" are similar but not identical to results from the "intervention period" of the SPRINT final report, which considered August 20, 2015 to July 29, 2016 as an observational post-intervention period [@sprintFinal]. The current study included August 20, 2015 to July 1, 2016 in the “trial phase” because antihypertensive medications continued to be provided until the final closeout visit on July 1, 2016. Also, as the NDI is continuously updated, the current analysis includes some deaths in 2016 that were not identified at the time of the final SPRINT report.

The Strategy of Blood Pressure Intervention in the Elderly Hypertensive Patients (STEP) trial enrolled 8511 Chinese patients 60 to 80 years of age with hypertension and randomized patients to a SBP target of 110 to less than 130 mm Hg (intensive treatment) or a target of 130 to less than 150 mm Hg (standard treatment) [@zhang_trial_2021]. STEP found a HR of 0.72 (95% CI 0.39, 1.32) with intensive versus standard treatment for CVD mortality after a median follow-up of 3.34 years, but did not find evidence of a benefit for all-cause mortality. In the current study, the protective effect of the SPRINT intensive treatment for all-cause mortality was attenuated several years before attenuation of the protective effect for CVD mortality. These results in combination with findings from the STEP trial suggest weaker evidence for reduced all-cause versus CVD mortality risk with intensive BP control.

Previous studies have found rising BP levels among US adults during the time period of the current study. General population studies of adults living in the US with hypertension found that the prevalence of uncontrolled BP (SBP ≥ 140 mm Hg or DBP ≥ 90 mm Hg) increased from 2013 to 2017 [@muntner2020htn]. In addition, an analysis of data from 464,585 adults enrolled in a Quest Diagnostics wellness program found SBP was between 1 mm Hg and 3 mm Hg higher, depending on age group and sex, in April through December of 2020 versus their corresponding values throughout 2019 [@laffin_rise_2022]. The current study shows that even for adults who have maintained intense SBP control for 3 years, increasing SBP levels may quickly diminish the protective effect. Combined with previous findings on rising BP levels among US adults, data from the current study emphasize the need for implementation of population- and community-level strategies to improve BP control in the US.

@@

This study has several limitations. First, while we restricted analyses to high quality NDI matches, misclassification in linking participants to the NDI is possible. Second, while several studies have shown reasonable performance of using NDI diagnosis codes for defining CVD mortality [@regards_ndi], it is not as robust as the adjudication process used in the primary follow-up for the trial. Third, information about SBP control after the trial was limited to routine outpatient SBP values in a subgroup of the SPRINT participants and was extracted from their EHR record, which is known to poorly reflect the standardized BP measurement process used during the trial [@drawz_concordance_2020]. While this prohibits definite conclusions about the absolute level of BP in both treatment groups, as well as pin-pointing when the between-group difference may have completely attenuated, the observation of steadily increasing SBP for participants in the intensive treatment and relatively stable SBP in the standard treatment group following the trial is likely still valid.

In conclusion, while intensive treatment produced beneficial effects on mortality during the trial, there was no evidence that this produced sustained benefits on cardiovascular and all-cause mortality subsequent to discontinuing the intervention protocol. Given steadily increasing mean SBP levels in participants randomized to intensive treatment after the trial, these results suggest that maintaining more intensive BP targets throughout adulthood will likely be essential for long-term CVD risk management.
